Edge sign prediction in general graphs and connection with Correlation Clustering

### 2017-08-09 cc_approx.tex
one more approximation algo and some tiny changes

In complete graphs, \textcite[Section 3]{Charikar2003} also give an improved $4$-approximation to
\mind{}, by rounding the same LP and using its solution in randomized algorithm. We will not
describe it in detail since a similar idea was used by \textcite{CCPivotConf05} with a better
approximation. To explain it, we first describe their randomized combinatorial algorithm \ccpivot{},
which gives a $3$-approximation of \mind{} on complete unweighted graphs (it has later been
derandomized while preserving its approximation guarantee~\autocite{derandomCCPivot08}). At each
iteration, we pick a node $u$ \uar{} and we create a cluster containing $u$ and all its neighbors
linked by a positive edges. On weighted complete graphs, they tweak this algorithm by using the
solution of the  LP~\eqref{eq:mindLP} to obtain different approximation factor depending on the
constraints imposed on the weight.\Todo{Comment on those constraints, especially the triangular ones
in the context of consensus clustering.} Recall that in the general formulation of the problem, each
edge carries two positive numbers: $w_{u,v}^+$ and $w_{u,v}^-$. If the weights respect the
probability constraints stating that for all edge $(u,v)$ in $E$, $w_{u,v}^+ + w_{u,v}^- = 1$, they
get a $2.5$-approximation. Note that unweighted graphs naturally fit into that case. If the weights
additionally respect the triangular inequality constraints stating that $w_{u,v}^- \leq w_{u,w}^- +
w_{w,v}^-$, this become a $2$-approximation. After solving the LP~\eqref{eq:mindLP} with additional
probability constraints, when picking a node $u$, each of its neighbors $v$ is added to the cluster
of $u$ with probability $x_{uv}$. \Textcite{Chawla2014} improve these two factors to respectively
$2.06$ and $1.5$ by exploiting the same idea but setting the probability to include each neighbor
$v$ of $u$ in the cluster of $u$ to be $1-f^+(x_{uv})$ if $(u,v)\in E^+$ and $1-f^-(x_{uv})$ if
$(u,v)\in E^-$, with a careful choice of $f^+$ and $f^-$. They also give a derandomized version of
their algorithm in the full version of the paper~\autocite[Theorem 23]{ChawlaArxiv14}.

@@

This conclude the presentation of the known approximation results on \pcc{}, that we summarize in
\autoref{tab:cc_approx}. As mentioned earlier, not having to set the number of clusters is an
attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.
The problem was studied by \textcite{Giotis2006} on general graph and we compiled their results in
\autoref{tab:cc_fixed}. On general graph and with $k$ being the number of clusters, 
they provide PTAS for \maxa{} running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and
for \mind{} running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}k^6\log d\right)}$.

>>>>

In complete graphs, \textcite[Section 3]{Charikar2003} also give an improved $4$-approximation to
\mind{}, by rounding the same LP and using its solution in randomized algorithm. We will not
describe it in detail since a similar idea was used by \textcite{CCPivotConf05} with a better
approximation. To explain it, we first describe their randomized combinatorial algorithm \ccpivot{},
which gives a $3$-approximation of \mind{} on complete unweighted graphs (it has later been
derandomized while preserving its approximation guarantee~\autocite{derandomCCPivot08}). At each
iteration, we pick a node $u$ \uar{} (called the pivot) and we create a cluster containing $u$ and all its neighbors
linked by a positive edges. On weighted complete graphs, they tweak this algorithm by using the
solution of the  LP~\eqref{eq:mindLP} to obtain different approximation factor depending on the
constraints imposed on the weight.\Todo{Comment on those constraints, especially the triangular ones
in the context of consensus clustering.} Recall that in the general formulation of the problem, each
edge carries two positive numbers: $w_{u,v}^+$ and $w_{u,v}^-$. If the weights respect the
probability constraints stating that for all edge $(u,v)$ in $E$, $w_{u,v}^+ + w_{u,v}^- = 1$, they
get a $2.5$-approximation. Note that unweighted graphs naturally fit into that case. If the weights
additionally respect the triangular inequality constraints stating that $w_{u,v}^- \leq w_{u,w}^- +
w_{w,v}^-$, this become a $2$-approximation. After solving the LP~\eqref{eq:mindLP} with additional
probability constraints, when picking a node $u$, each of its neighbors $v$ is added to the cluster
of $u$ with probability $x_{uv}$. \Textcite{Chawla2014} improve these two factors to respectively
$2.06$ and $1.5$ by exploiting the same idea but setting the probability to include each neighbor
$v$ of $u$ in the cluster of $u$ to be $1-f^+(x_{uv})$ if $(u,v)\in E^+$ and $1-f^-(x_{uv})$ if
$(u,v)\in E^-$, with a careful choice of $f^+$ and $f^-$. They also give a derandomized version of
their algorithm in the full version of the paper~\autocite[Theorem 23]{ChawlaArxiv14}.

@@

This conclude the presentation of the known approximation results on \pcc{}, that we summarize in
\autoref{tab:cc_approx}. As mentioned earlier, not having to set the number of clusters is an
attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.  The
problem was studied by \textcite{Giotis2006} on general graph and we compiled their results in
\autoref{tab:cc_fixed}. On complete unweighted graph and with $k$ being the number of clusters, they
provide PTAS for \maxa{}$_k$ running in $nk^{O(\epsilon^{-3}\log(\frac{k}{\epsilon}))}$ time and for
\mind{}$_k$ running in $n^{O\left(\epsilon^{-2} 9^k\right)}\log(n)$ time. The latter was improved by
\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}k^6\log d\right)}$
and that can handle weighted graphs. For $k=2$, \mind$_2$ admits a faster local search method with a
factor $2$ approximation~\autocite{Coleman2008}.
For complete weighted graph, \textcite{WeightedMaxAPTAS08} provide a PTAS for
\maxa{}$_k$ under the condition that the ratio between the largest and smallest weights is bounded by a
constant.

### 2017-08-09 cc_sign_relation.tex
one more approximation algo and some tiny changes

In real data though, we do not expect neither strong nor weak balance to hold, for they are fairly
demanding model. Indeed, three of the real networks considered in \autoref{chap:troll} have been
repeatedly shown to be unbalanced, although the extent of this unbalance depends of the importance
given to longer cycle compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
{measureUnbalance17} For the same reason, in virtually any real instance of the \pcc{} problem, the
optimal solution will incur some disagreements. \Textcite[Section 5.5]{BookKleinberg2010}
nonetheless prove that even when only a fraction of the triangle are strongly balanced in a complete
graph, the two clusters structure is still present, although it does not cover the whole graph
anymore. Formally statement
The proof makes essential use of the strong balance (which is ok since next section also deal with
2-CC) and the completeness of the graph (which is a far less realistic assumption) yet it would be
interesting to see if it can be extended.

@@

there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration index)
which they claimed has been peer reviewed? They also mention their method to compute frustration
exactly \url{https://arxiv.org/abs/1611.09030} along with existing work (for $n=50$)
\url{https://doi.org/10.1016/j.socnet.2013.09.002} (leading to an 1996 heuristics for \pcc{} in
section 3 of \url{https://doi.org/10.1016/0378-8733(95)00259-6})

>>>>

In real data though, we do not expect neither strong nor weak balance to hold, for they are fairly
demanding model. Indeed, three of the real networks considered in \autoref{chap:troll} have been
repeatedly shown to be unbalanced, although the extent of this unbalance depends of the importance
given to longer cycle compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
{measureUnbalance17} For the same reason, in virtually any real instance of the \pcc{} problem, the%
% there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration
% index) which they claimed has been peer reviewed?
optimal solution will incur some disagreements. \Textcite[Section 5.5]{BookKleinberg2010}
nonetheless prove that even when only a fraction of the triangle are strongly balanced in a complete
graph, the two clusters structure is still present, although it does not cover the whole graph
anymore. Formally statement
The proof makes essential use of the strong balance (which is ok since next section also deal with
2-CC) and the completeness of the graph (which is a far less realistic assumption) yet it would be
interesting to see if it can be extended.

@@
